Monitoring misinformation related interventions by Facebook, Twitter and YouTube: methods and illustration.

### 2021-08-23 article/draft.tex
write last paragraph of discussion

A last concern regards the growing restrictions to access APIs \cite{api}.
In April 2018, likely as a response to the Cambridge Analytica scandal, Facebook severely limited access to their official APIs\footnote{\url{https://about.fb.com/news/2018/04/restricting-data-access/}, \url{http://thepoliticsofsystems.net/2018/08/facebooks-app-review-and-how-independent-research-just-got-a-lot-harder/}}, 
and CrowdTangle is today a reliable source for researchers to access Facebook data. 
However, it should be noted that CrowdTangle has been recently used by journalists to show that Facebook is spreading biased information\footnote{\url{https://www.economist.com/graphic-detail/2020/09/10/facebook-offers-a-distorted-view-of-american-news}, \url{https://twitter.com/facebookstop10}. 
The last link is a Twitter account showing each day the ten most shared link on Facebook, most of them coming from extreme right and misinformation sources. 
Facebook has reacted to this here: \url{https://about.fb.com/news/2020/11/what-do-people-actually-see-on-facebook-in-the-us/}}, 
and a general suspicion is growing that its access might be shut down in a near future to journalists and researchers.\footnote{\url{https://www.nytimes.com/2021/07/14/technology/facebook-data.html}, \url{https://www.bloomberg.com/news/articles/2021-08-03/facebook-disables-accounts-tied-to-nyu-research-project}}
Regarding Twitter and YouTube, it is still possible and relatively easy to access their official APIs, although the number of Tweets that can be retrieved has decreased between the Twitter API v1 and v2. 
In the v2, there is a monthly cap of 10,000,000 Tweets for the academic research track\footnote{https://developer.twitter.com/en/docs/projects/overview},
while there was no monthly limit in the v1, only a limit in the number of requests that could be made in a 15-minute window\footnote{https://developer.twitter.com/en/docs/twitter-api/v1/rate-limits}.
We thus noticed that we could get around 10 times more Tweets per month with the v1 when compared to the v2 for one specific collection.

@@

{\color{brown} Tackling online misinformation is an active area of research and adapted interventions might be required to avoid undesired effects. 
psychological and indirect effects (Pennycook, etc.). 
How labels can work against their intent.}

>>>>

A last concern regards the growing restrictions to access APIs (Perriam et al. 2019~\cite{api}).
In April 2018, likely as a response to the Cambridge Analytica scandal, Facebook severely limited access to their official APIs\footnote{\url{https://about.fb.com/news/2018/04/restricting-data-access/}, \url{http://thepoliticsofsystems.net/2018/08/facebooks-app-review-and-how-independent-research-just-got-a-lot-harder/}}, 
and CrowdTangle is today a reliable source for researchers to access Facebook data. 
However, it should be noted that CrowdTangle has been recently used by journalists to show that Facebook is spreading biased information\footnote{\url{https://www.economist.com/graphic-detail/2020/09/10/facebook-offers-a-distorted-view-of-american-news}, \url{https://twitter.com/facebookstop10}. 
The last link is a Twitter account showing each day the ten most shared link on Facebook, most of them coming from extreme right and misinformation sources. 
Facebook has reacted to this here: \url{https://about.fb.com/news/2020/11/what-do-people-actually-see-on-facebook-in-the-us/}}, 
and a general suspicion is growing that its access might be shut down in a near future to journalists and researchers.\footnote{\url{https://www.nytimes.com/2021/07/14/technology/facebook-data.html}, \url{https://www.bloomberg.com/news/articles/2021-08-03/facebook-disables-accounts-tied-to-nyu-research-project}}
Regarding Twitter and YouTube, it is still possible and relatively easy to access their official APIs, although the number of Tweets that can be retrieved has decreased between the Twitter API v1 and v2. 
In the v2, there is a monthly cap of 10,000,000 Tweets for the academic research track\footnote{https://developer.twitter.com/en/docs/projects/overview},
while there was no monthly limit in the v1, only a limit in the number of requests that could be made in a 15-minute window\footnote{https://developer.twitter.com/en/docs/twitter-api/v1/rate-limits}.
We thus noticed that we could get around 10 times more Tweets per month with the v1 when compared to the v2 for one specific collection.

@@

How to best tackle online misinformation is an active area of research, and developing effective interventions against misinformation depends on understanding the psychology of interacting with social media (Pennycook et Rand 2021~\cite{pennycook2021psychology}). 
Adding fact-checking labels to controversial content is certainly the intervention whose effectiveness was the most studied so far. 
It was shown that credibility indicators do reduce misperceptions and sharing, although their impact varied vastly depending on the presentation of the warnings (Yaqub 2020~\cite{yaqub2020effects}, Sharevski 2021~\cite{sharevski2021misinformation}).
Moreover, a fact-checking label can lead to unpredicted consequences depending on the perception of the person reading it.
On Facebook, Twitter and YouTube, fact-checking messages are currently applied only to some content, while the rest of the platform remains unlabelled. 
Such sparce application of warnings can lead to an `implied truth' effect where users may assume that content without warnings have actually been verified and shown to be true (Pennycook et al. 2020~\cite{pennycook2020implied}).
Using another approach, Mosleh et al. (\cite{mosleh2021perverse}) identified Twitter users sharing false news and replied to their false tweets with links to fact-checking websites.
They observed a decrease in quality and accuracy of the content subsequently shared by the corrected user. 
These works highlight the necessity to investigate potential unintended effects of well-meaning interventions.
The methods presented in this paper can help to monitor not only the platforms' actions against misinformations, but also their subsequent consequences in real-life settings.
